### 2019-06-28 doc/fast_ssa/paper.tex
working on the documentation

%\documentclass[manuscript,endfloat]{geophysics}
\documentclass[revised,endfloat]{geophysics}
%\documentclass[twocolumn,endfloat]{geophysics}

\usepackage{xy}
\usepackage{natbib}
\usepackage{graphicx}
\usepackage{url}
\usepackage{rotating}
\usepackage{lscape}
\usepackage{color}
\usepackage{rotating}
\usepackage{amsmath}
\usepackage{mathrsfs}
\usepackage{amssymb}      
\usepackage{algpseudocode}
\usepackage{booktabs}
 \usepackage{tikz}
 \usetikzlibrary{matrix}

\usepackage[ruled,linesnumbered]{algorithm2e}
\SetKwProg{Fn}{Function}{}{}

\usepackage{mathtools}
\DeclarePairedDelimiter\ceil{\lceil}{\rceil}
\DeclarePairedDelimiter\floor{\lfloor}{\rfloor}

\setlength\parindent{0pt}

\setlength\heavyrulewidth{0.25ex}

\DeclareGraphicsExtensions{.pdf}
\DeclareMathAlphabet\mathbfcal{OMS}{cmsy}{b}{n}
\DeclareMathOperator*{\argmin}{arg\,min}

\begin{document}

\title{Structured matrix}
\renewcommand{\thefootnote}{\fnsymbol{footnote}}

\address{
\footnotemark[1]Department of Physics,\\
University of Alberta, \\
Edmonton, Alberta, Canada. \\}

\author{Wenlei Gao\footnotemark[1]}

\maketitle

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% ABSTRACT                                                                   %
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{abstract}
The documentation for fast MSSA and relationship between multi-dimensional convolution and structured matrix, such as Circulant, Toeplitz and Hankel matrix.
\end{abstract}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}
\citep{korobeynikov2009} proposed the computation and memory efficient method for MSSA.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Method}
We start from the definition of discrete fourier transform.

\subsection{Discrete Fourier transform (DFT)}
For a vector ${\bf s}$ length of $N$, the discrete Fourier transform is  defined as
\begin{equation}
\tilde{s}_k  = \sum_{l=0}^{N-1} e^{-i 2 \pi k l / N} s_l \,,
\label{eq1}
\end{equation}
where the index $k = 0, 1, 2, \cdots, N-1$. Suppose the vector ${\bf s}$ represent a time series length of $N$ with time sampling interval is $\Delta t$. So the frequency sampling interval $\Delta f$ can be computed as 
\begin{equation}
\Delta f  = 1/ \Delta t / N = \frac{1}{N \Delta t}
\label{eq2}
\end{equation}
The Fourier transform for this time series in continueous case can be expressed as
\begin{equation}
\tilde{s}(f)  = \int e^{-i 2 \pi f t} s(t) dt 
\label{eq3}
\end{equation}
In discrete case and assume the frequency $f$ can be expressed as $f = k \Delta f$ and using the expression in equation \ref{eq2}
\begin{equation}
\begin{split}
\tilde{s}_{k \Delta f}  &= \sum_{l=0}^{N-1} e^{-i 2 \pi \, k \Delta f \,  l \Delta t} s_{l \Delta t}  \\
\tilde{s}_{k \Delta f}  &= \sum_{l=0}^{N-1} e^{-i 2 \pi \, k \frac{1}{N \Delta t} \,  l \Delta t} s_{l \Delta t} \\
\tilde{s}_{k}  &= \sum_{l=0}^{N-1} e^{-i 2 \pi \, k l / N} s_l \,.
\end{split}
\label{eq4}
\end{equation}
The the last equation is exactly same as the equation \ref{eq1}. Through this derivation, we can understand the physical meaning of the Fourier coefficients obtained via discrete Fourier transform. Note that the first coefficient corresponding to $0$ frequency and so as to the time series. The first element of ${\bf s}$ is the sample at $0$ time.

The above operation can be expressed into the matrix-vector form. To simplify the definition of Fourier transform matrix, we define a variable $z = e^{-i \pi  / N}$, so the Fourier transform matrix, which is a $N$-by-$N$ square matrix and the element $F_{kl} = z^{kl}$. It can be expanded as
\begin{equation}
{\bf F}_N = \begin{bmatrix}
1 & 1 & \cdots & 1 \\
1 & z & \cdots & z^{N-1} \\
1 & z^2 & \cdots & z^{2(N-1)} \\
\vdots & \vdots & \ddots & \vdots \\
1 & z^{N-1} & \cdots & z^{(N-1)(N-1)} \\
\end{bmatrix}
\label{eq5}
\end{equation}
As ${\bf F}_N$ is a orthogonal (but not orthonormal) matrix, we can be easily get the its inverse matrix
\begin{equation}
{\bf F}_N^{-1} = \frac{1}{N} {\bf F}_N^H = \frac{1}{N} {\bf F}_N^{\ast} \,
\label{eq6}
\end{equation}
where the super-script $^H$ represent complex conjugate transpose and $\ast$ indicate conjugate transpose. As the matrix ${\bf F}_N$ is a symmetrical matrix, that's why we can get equation \ref{eq6}.

\subsection{Circulant matrix}
An $9 \times 9$ circulant matrix is defined as
\begin{equation}
{\bf C} = 
\begin{bmatrix}
c_1 & c_9 & c_8 & c_7 & c_6 & c_5 &  c_4 & c_3 & c_2 \\ 
c_2 & c_1 & c_9 & c_8 & c_7 & c_6 &  c_5 & c_4 & c_3 \\ 
c_3 & c_2 & c_1 & c_9 & c_8 & c_7 &  c_6 & c_5 & c_4 \\ 
c_4 & c_3 & c_2 & c_1 & c_9 & c_8 &  c_7 & c_6 & c_5 \\ 
c_5 & c_4 & c_3 & c_2 & c_1 & c_9 &  c_8 & c_7 & c_6 \\ 
c_6 & c_5 & c_4 & c_3 & c_2 & c_1 &  c_9 & c_8 & c_7 \\ 
c_7 & c_6 & c_5 & c_4 & c_3 & c_2 &  c_1 & c_9 & c_8 \\ 
c_8 & c_7 & c_6 & c_5 & c_4 & c_3 &  c_2 & c_1 & c_9 \\ 
c_9 & c_8 & c_7 & c_6 & c_5 & c_4 &  c_3 & c_2 & c_1 \\ 
\end{bmatrix}
\end{equation} 
A circulant matrix is fully specified by its first column ${\bf c} = \begin{bmatrix}
c_1 & c_2 &  c_3 &  c_4 &  c_5 &  c_6 &  c_7 &  c_8 &  c_9  \end{bmatrix}^T$, we know that the product between circulant matrix and a vector ${\bf v}$ length of $N$ is equivalent to the circular convolution between vector ${\bf c}$ and ${\bf v}$, and the circular convolution can be computed efficiently via Fast Fourier transform (FFT). So we have
\begin{equation}
{\bf Cv} = {\bf F}_N^{-1} \left(  {\bf F}_N {\bf c} \circ {\bf F}_N {\bf v} \right)  =  {\bf F}_N^{-1} \left(  diag({\bf F}_N {\bf c}) \cdot {\bf F}_N {\bf v} \right) \,,
\end{equation}
where the symbol $\circ$ and $diag$ represents Hardmard (element-wise) multiplication and building diagonal matrix from a vector.

\subsection{Toeplitz matrix}
Toeplitz matrix is associated with linear convolution, suppose we have two vectors ${\bf s} = \begin{bmatrix}  s_1 & s_2 & s_3 & s_4 & s_5  \end{bmatrix}^T$ and ${\bf f} =  \begin{bmatrix}  f_1 & f_2 & f_3 & f_4 & f_5  \end{bmatrix}^T$,  the linear convolution of these two vectors is vector ${\bf r}$ length of $9$. We can represent the convolution into matrix-vector form as
\begin{equation}
{\bf r} = {\bf  h} \ast {\bf f} = \begin{bmatrix}
r_1 \\
r_2 \\
r_3 \\
r_4 \\
r_5 \\
r_6 \\
r_7 \\
r_8 \\
r_9 \\
\end{bmatrix}
= \begin{bmatrix}
s_1    & \quad & \quad & \quad & \quad  \\ 
s_2    & s_1    & \quad & \quad & \quad  \\ 
s_3    & s_2    & s_1 & \quad & \quad \\ 
s_4    & s_3 & s_2 & s_1 & \quad  \\ 
s_5    & s_4 & s_3 & s_2 & s_1  \\ 
\quad & s_5 & s_4 & s_3 & s_2 \\ 
\quad & \quad & s_5 & s_4 & s_3  \\ 
\quad & \quad & \quad & s_5 & s_4 \\ 
\quad & \quad & \quad & \quad & s_5  \\ 
\end{bmatrix}
\begin{bmatrix}
f_1 \\
f_2 \\
f_3 \\
f_4 \\
f_5 \\
\end{bmatrix}
\end{equation}
The efficient linear convolution usually padding $0$ to vectors make their length equal to $9$, after padding zeros, we get new vector $\hat{\bf s} = \begin{bmatrix}  s_1 & s_2 & s_3 & s_4 & s_5 & 0 & 0 & 0 & 0 \end{bmatrix}$ and  $\hat{\bf f} = \begin{bmatrix}  f_1 & f_2 & f_3 & f_4 & f_5 & 0 & 0 & 0 & 0 \end{bmatrix} $, the circular convolution between $\hat{\bf s}$ and $\hat{\bf f}$ can be expressed as
\begin{equation}
\hat{\bf r} = \hat{\bf s} \circledast \hat{\bf f} = \begin{bmatrix}
s_1 & 0     & 0     & 0    & 0     & s_5 &  s_4 & s_3 & s_2 \\ 
s_2 & s_1 & 0     & 0    & 0     & 0     &  s_5 & s_4 & s_3 \\ 
s_3 & s_2 & s_1 & 0    & 0     & 0     &  0    & s_5 & s_4 \\ 
s_4 & s_3 & s_2 & s_1 & 0     & 0     &  0 & 0 & s_5 \\ 
s_5 & s_4 & s_3 & s_2 & s_1 & 0     &  0 & 0 & 0 \\ 
0    & s_5  & s_4 & s_3 & s_2 & s_1 &  0 & 0 & 0 \\ 
0    & 0     & s_5 & s_4  & s_3 & s_2 &  s_1 & 0 & 0 \\ 
0    & 0     & 0    & s_5  & s_4 & s_3 &  s_2 & s_1 & 0 \\ 
0    & 0     & 0    & 0     &  s_5 & s_4 &  s_3 & s_2 & s_1 \\ 
\end{bmatrix}
\begin{bmatrix}
f_1 \\
f_2 \\
f_3 \\
f_4 \\
f_5 \\
0 \\
0 \\
0 \\
0 \\
\end{bmatrix}
\label{eq10}
\end{equation}
When can see that $\hat{\bf r} = {\bf r}$, so linear convolution can be efficiently computed by padding zeros to vectors then compute circular convolution in frequency domain.

Let's consider another Toeplitz matrix built from a vector ${\bf c} = \begin{bmatrix}
c_1 & c_2 & c_3 & c_4 & c_5 & c_6 & c_7 & c_8 & c_9 \end{bmatrix}$ with length $N=9$, we first determine two integer: $L$ and $K$ make $L+K-1= N $. To make the matrix close to square form, we determine them by
\begin{equation}
L = \floor*{\frac{N}{2}}+1 ,\,\, K = N+1-L
\label{eq11}
\end{equation}
in this special case, $N=9$, so $L=K=5$, the Toeplitz matrix build from this vector can be expressed as
\begin{equation}
{\bf T} = 
\begin{bmatrix}
c_5 & c_4 & c_3 & c_2 & c_1 \\ 
c_6 & c_5 & c_4 & c_3 & c_2 \\ 
c_7 & c_6 & c_5 & c_4 & c_3 \\ 
c_8 & c_7 & c_6 & c_5 & c_4 \\ 
c_9 & c_8 & c_7 & c_6 & c_5 \\ 
\end{bmatrix}
\label{eq12}
\end{equation} 
The vector ${\bf c}$ can be uniquely determined from the last column and last row. 
There are two way to embedding this Teoplitz matrix into a circular matrix, the first to embedding this matrix at the lower-left part of the circular matrix created by vector ${\bf c}$, which shows as below as
\begin{center}
\begin{tikzpicture}
  \matrix (m)[
    matrix of math nodes,
    nodes in empty cells,
    minimum width=width("123"),
  ] {
c_1 & c_9 & c_8 & c_7 & c_6 & c_5 &  c_4 & c_3 & c_2 \\ 
c_2 & c_1 & c_9 & c_8 & c_7 & c_6 &  c_5 & c_4 & c_3 \\ 
c_3 & c_2 & c_1 & c_9 & c_8 & c_7 &  c_6 & c_5 & c_4 \\ 
c_4 & c_3 & c_2 & c_1 & c_9 & c_8 &  c_7 & c_6 & c_5 \\ 
c_5 & c_4 & c_3 & c_2 & c_1 & c_9 &  c_8 & c_7 & c_6 \\ 
c_6 & c_5 & c_4 & c_3 & c_2 & c_1 &  c_9 & c_8 & c_7 \\ 
c_7 & c_6 & c_5 & c_4 & c_3 & c_2 &  c_1 & c_9 & c_8 \\ 
c_8 & c_7 & c_6 & c_5 & c_4 & c_3 &  c_2 & c_1 & c_9 \\ 
c_9 & c_8 & c_7 & c_6 & c_5 & c_4 &  c_3 & c_2 & c_1 \\ 
  } ;
\draw (m-9-1.south west) rectangle (m-5-5.north east);
\end{tikzpicture}
\end{center}
The above circulant matrix is called ${\bf C}_1$. We can create a new vector $\hat{\bf c} = \begin{bmatrix} c_5 & c_6 & c_7 & c_8 & c_9 & c_1 & c_2 & c_3 & c_4 \end{bmatrix}^T$, which is obtained by permutating the elements of the original vector, the Toeplitz matrix can be embedded at the upper-left part of the circulant matrix created from $\hat{\bf c}$ shows as bellow
 \begin{center}
\begin{tikzpicture}
  \matrix (m)[
    matrix of math nodes,
    nodes in empty cells,
    minimum width=width("123"),
  ] {
c_5 & c_4 & c_3 & c_2 & c_1 & c_9 &  c_8 & c_7 & c_6 \\ 
c_6 & c_5 & c_4 & c_3 & c_2 & c_1 &  c_9 & c_8 & c_7 \\ 
c_7 & c_6 & c_5 & c_4 & c_3 & c_2 &  c_1 & c_9 & c_8 \\ 
c_8 & c_7 & c_6 & c_5 & c_4 & c_3 &  c_2 & c_1 & c_9 \\ 
c_9 & c_8 & c_7 & c_6 & c_5 & c_4 &  c_3 & c_2 & c_1 \\ 
c_1 & c_9 & c_8 & c_7 & c_6 & c_5 &  c_4 & c_3 & c_2 \\ 
c_2 & c_1 & c_9 & c_8 & c_7 & c_6 &  c_5 & c_4 & c_3 \\ 
c_3 & c_2 & c_1 & c_9 & c_8 & c_7 &  c_6 & c_5 & c_4 \\ 
c_4 & c_3 & c_2 & c_1 & c_9 & c_8 &  c_7 & c_6 & c_5 \\ 
} ;
\draw (m-5-1.south west) rectangle (m-1-5.north east);
\end{tikzpicture}
\end{center}
 and call this new matrix as ${\bf C}_2$.

Suppose we need to compute the product between Toeplitz matrix ${\bf T}$ and a vector ${\bf v}$
\begin{equation}
{\bf r} = {\bf T}{\bf v} \,
\end{equation} 
where the length of ${\bf v}$ is $K$.
There are two way to do it efficiently via FFT, and these two ways corresponding to the two embedding methods. The first step is to padding vector ${\bf v}$ with $L-1$ $0$s get a new vector $\hat{\bf v}$. Following above example, $K=L=5$ and ${\bf v} = \begin{bmatrix} v_1 & v_2 & v_3 & v_4 & v_5  \end{bmatrix}^T$, and the padded vector $\hat{\bf v} = \begin{bmatrix} v_1 & v_2 & v_3 & v_4 & v_5 & 0 & 0 & 0 & 0 \end{bmatrix}^T$, The product between ${\bf C}_1$ and $\hat{\bf v}$ are
\begin{equation}
\hat{\bf r}_1 = {\bf C}_1 \hat{\bf v} = \begin{bmatrix}
c_1 & c_9 & c_8 & c_7 & c_6 & c_5 &  c_4 & c_3 & c_2 \\ 
c_2 & c_1 & c_9 & c_8 & c_7 & c_6 &  c_5 & c_4 & c_3 \\ 
c_3 & c_2 & c_1 & c_9 & c_8 & c_7 &  c_6 & c_5 & c_4 \\ 
c_4 & c_3 & c_2 & c_1 & c_9 & c_8 &  c_7 & c_6 & c_5 \\ 
c_5 & c_4 & c_3 & c_2 & c_1 & c_9 &  c_8 & c_7 & c_6 \\ 
c_6 & c_5 & c_4 & c_3 & c_2 & c_1 &  c_9 & c_8 & c_7 \\ 
c_7 & c_6 & c_5 & c_4 & c_3 & c_2 &  c_1 & c_9 & c_8 \\ 
c_8 & c_7 & c_6 & c_5 & c_4 & c_3 &  c_2 & c_1 & c_9 \\ 
c_9 & c_8 & c_7 & c_6 & c_5 & c_4 &  c_3 & c_2 & c_1 \\ 
\end{bmatrix}
\begin{bmatrix}
v_1 \\
v_2 \\
v_3 \\
v_4 \\
v_5 \\
0 \\
0 \\
0 \\
0 \\
\end{bmatrix}
\end{equation}
where the last $L$ elements of $\hat{\bf r}_1$ equal to ${\bf r}$. On the other hand, we can take advantage of the circulate matrix ${\bf C}_2$, we have
\begin{equation}
\hat{\bf r}_2 = {\bf C}_2 \hat{\bf v} = \begin{bmatrix}
c_5 & c_4 & c_3 & c_2 & c_1 & c_9 &  c_8 & c_7 & c_6 \\ 
c_6 & c_5 & c_4 & c_3 & c_2 & c_1 &  c_9 & c_8 & c_7 \\ 
c_7 & c_6 & c_5 & c_4 & c_3 & c_2 &  c_1 & c_9 & c_8 \\ 
c_8 & c_7 & c_6 & c_5 & c_4 & c_3 &  c_2 & c_1 & c_9 \\ 
c_9 & c_8 & c_7 & c_6 & c_5 & c_4 &  c_3 & c_2 & c_1 \\ 
c_1 & c_9 & c_8 & c_7 & c_6 & c_5 &  c_4 & c_3 & c_2 \\ 
c_2 & c_1 & c_9 & c_8 & c_7 & c_6 &  c_5 & c_4 & c_3 \\ 
c_3 & c_2 & c_1 & c_9 & c_8 & c_7 &  c_6 & c_5 & c_4 \\ 
c_4 & c_3 & c_2 & c_1 & c_9 & c_8 &  c_7 & c_6 & c_5 \\ 
\end{bmatrix}
\begin{bmatrix}
v_1 \\
v_2 \\
v_3 \\
v_4 \\
v_5 \\
0 \\
0 \\
0 \\
0 \\
\end{bmatrix}
\end{equation}
So the first $L$ elements of $\hat{\bf r}_2$ equal to ${\bf r}$. We summarize the efficient computation of the product between Toeplitz matrix and a vector in Algorithm 1 and Algorithm 2
\begin{algorithm}
      \SetAlgoLined
      \Fn{ ${\bf v}= ttv({\bf c}, {\bf v}) $}{
      length of ${\bf c}$: $N = length({\bf c})$   \\
      compute $L = \floor*{\frac{N}{2}}+1$, $K = N+1 - L$   \\
      padding $L-1$ $0$s to the end of ${\bf v}$ to get a new vector $\hat{\bf v}$ with length of $N$\\
      ${\bf r}_1 = {\bf F}_N^{-1} \left(   {\bf F}_N {\bf c} \circ {\bf F}_N  \hat{\bf v}\right) $ \\
      take the last $L$ elements from ${\bf r}_1$ 
      }
      \caption{Teoplitz-Times-Vector}
\end{algorithm}

Another method is summarized in Algorithm 2 as 
\begin{algorithm}
      \SetAlgoLined
      \Fn{ ${\bf v}= ttv({\bf c}, {\bf v}) $}{
      length of ${\bf c}$: $N = length({\bf c})$   \\
      compute $L = \floor*{\frac{N}{2}}+1$, $K = N+1 - L$   \\
      padding $L-1$ $0$s to the end of ${\bf v}$ to get a new vector $\hat{\bf v}$ with length of $N$\\
      move the first $K-1$ element of ${\bf c}$ to the end, get a new vector $\hat{\bf c}$
      ${\bf r}_2 = {\bf F}_N^{-1} \left(   {\bf F}_N \hat{\bf c} \circ {\bf F}_N  \hat{\bf v}\right) $ \\
      take the first $L$ elements from ${\bf r}_2$ 
      }
      \caption{Teoplitz-Times-Vector}
\end{algorithm}

\subsection{Hankel matrix}
Hankel matrix of size $L \times K$ created from the same vector ${\bf c}$ is given as
\begin{equation}
{\bf H} = 
\begin{bmatrix}
c_1 & c_2 & c_3 & c_4 & c_5 \\ 
c_2 & c_3 & c_4 & c_5 & c_6 \\ 
c_3 & c_4 & c_5 & c_6 & c_7 \\ 
c_4 & c_5 & c_6 & c_7 & c_8 \\ 
c_5 & c_6 & c_7 & c_8 & c_9 \\ 
\end{bmatrix}
\label{eq16}
\end{equation}
which can be determined from the first column and last row.

\textcolor{red}{Note: The circulant matrix and Teoplitz matrix are constant on the diagonals. However, Hankel matrix are constant on the anti-diagonals.}

One can convert a Hankel matrix to a Toeplitz matrix by reversing its columns. The columns reversing process can be represented by right multiplying a $K \times k$ matrix ${\bf R}$, whose elements are $1$ on the main anti-diagonals and 0 elsewhere. Follow the toy example, we can get
\begin{equation}
{\bf T}= \begin{bmatrix}
c_5 & c_4 & c_3 & c_2 & c_1 \\ 
c_6 & c_5 & c_4 & c_3 & c_2 \\ 
c_7 & c_6 & c_5 & c_4 & c_3 \\ 
c_8 & c_7 & c_6 & c_5 & c_4 \\ 
c_9 & c_8 & c_7 & c_6 & c_5 \\ 
\end{bmatrix} = {\bf HR} = 
\begin{bmatrix}
c_1 & c_2 & c_3 & c_4 & c_5 \\ 
c_2 & c_3 & c_4 & c_5 & c_6 \\ 
c_3 & c_4 & c_5 & c_6 & c_7 \\ 
c_4 & c_5 & c_6 & c_7 & c_8 \\ 
c_5 & c_6 & c_7 & c_8 & c_9 \\ 
\end{bmatrix}
\begin{bmatrix}
\quad & \quad &  \quad &  \quad & 1 \\
\quad & \quad &  \quad &  1 & \quad \\
\quad & \quad &  1 &  \quad & \quad \\
\quad & 1 &  \quad &  \quad & \quad \\
1 & \quad &  \quad &  \quad & \quad \\
\end{bmatrix}
\end{equation}

Before we discuss about the multiplication between Hankel matrix and a vector, we first look at the property of the column reversing matrix ${\bf R}$, we have 
\begin{equation}
{\bf R} = {\bf R}^T = {\bf R}^{-1}
\end{equation}
So the product between Hankel matrix ${\bf H}$ and a vector ${\bf v}$ is given as 
\begin{equation}
{\bf r} = {\bf Hv} = {\bf HRR}^{-1}{\bf v} = {\bf HRRv} = ({\bf HR})({\bf Rv}) = {\bf T} \hat {\bf v}
\end{equation}
where $\bf{v}$ is the reversed version of the vector ${\bf v}$.

\bibliographystyle{seg}
\bibliography{paper}

\end{document}